Multiresolution Neural Field Estimation

### 2011-02-28 Multiresolution_framework/ltx/manuscript.tex
deleted some of the top matter junk, mostly to see if I can push to the repo
been ruthless with the first paragraph of the state and parameter estimation bit

%% bare_jrnl.tex
%% V1.3
%% 2007/01/11
%% by Michael Shell
%% see http://www.michaelshell.org/
%% for current contact information.
%%
%% This is a skeleton file demonstrating the use of IEEEtran.cls
%% (requires IEEEtran.cls version 1.7 or later) with an IEEE journal paper.
%%
%% Support sites:
%% http://www.michaelshell.org/tex/ieeetran/
%% http://www.ctan.org/tex-archive/macros/latex/contrib/IEEEtran/
%% and
%% http://www.ieee.org/

% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
% *** with production work. IEEE's font choices can trigger bugs that do  ***
% *** not appear when using other class files.                            ***
% The testflow support page is at:
% http://www.michaelshell.org/tex/testflow/

%%*************************************************************************
%% Legal Notice:
%% This code is offered as-is without any warranty either expressed or
%% implied; without even the implied warranty of MERCHANTABILITY or
%% FITNESS FOR A PARTICULAR PURPOSE! 
%% User assumes all risk.
%% In no event shall IEEE or any contributor to this code be liable for
%% any damages or losses, including, but not limited to, incidental,
%% consequential, or any other damages, resulting from the use or misuse
%% of any information contained here.
%%
%% All comments are the opinions of their respective authors and are not
%% necessarily endorsed by the IEEE.
%%
%% This work is distributed under the LaTeX Project Public License (LPPL)
%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
%% distributed and modified. A copy of the LPPL, version 1.3, is included
%% in the base LaTeX documentation of all distributions of LaTeX released
%% 2003/12/01 or later.
%% Retain all contribution notices and credits.
%% ** Modified files should be clearly indicated as such, including  **
%% ** renaming them and changing author support contact information. **
%%
%% File list of work: IEEEtran.cls, IEEEtran_HOWTO.pdf, bare_adv.tex,
%%                    bare_conf.tex, bare_jrnl.tex, bare_jrnl_compsoc.tex
%%*************************************************************************

% Note that the a4paper option is mainly intended so that authors in
% countries using A4 can easily print to A4 and see how their papers will
% look in print - the typesetting of the document will not typically be
% affected with changes in paper size (but the bottom and side margins will).
% Use the testflow package mentioned above to verify correct handling of
% both paper sizes by the user's LaTeX system.
%
% Also note that the "draftcls" or "draftclsnofoot", not "draft", option
% should be used if it is desired that the figures are to be displayed in
% draft mode.

@@

% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[journal]{../sty/IEEEtran}

% Some very useful LaTeX packages include:
% (uncomment the ones you want to load)

% *** MISC UTILITY PACKAGES ***
%
%\usepackage{ifpdf}
% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
% compilation based on whether the output is pdf or dvi.
% usage:
% \ifpdf
%   % pdf code
% \else
%   % dvi code
% \fi
% The latest version of ifpdf.sty can be obtained from:
% http://www.ctan.org/tex-archive/macros/latex/contrib/oberdiek/
% Also, note that IEEEtran.cls V1.7 and later provides a builtin
% \ifCLASSINFOpdf conditional that works the same way.
% When switching from latex to pdflatex and vice-versa, the compiler may
% have to be run twice to clear warning/error messages.

% *** CITATION PACKAGES ***
%
\usepackage{cite}
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
% \cite{} output to follow that of IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
% \cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 4.0 (2003-05-27) and later if using hyperref.sty. cite.sty does
% not currently provide for hyperlinked citations.
% The latest version can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/cite/
% The documentation is contained in the cite.sty file itself.

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
   \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../pdf/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
\fi
% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation can
% be obtained at: 
% http://www.ctan.org/tex-archive/macros/latex/required/graphics/
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found as epslatex.ps or
% epslatex.pdf at: http://www.ctan.org/tex-archive/info/
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex

% *** MATH PACKAGES ***
%
\usepackage[cmex10]{amsmath}
\interdisplaylinepenalty=2500
% A popular package from the American Mathematical Society that provides
% many useful and powerful commands for dealing with mathematics. If using
% it, be sure to load this package with the cmex10 option to ensure that
% only type 1 fonts will utilized at all point sizes. Without this option,
% it is possible that some math symbols, particularly those within
% footnotes, will be rendered in bitmap form which will result in a
% document that can not be IEEE Xplore compliant!
%
% Also, note that the amsmath package sets \interdisplaylinepenalty to 10000
% thus preventing page breaks from occurring within multiline equations. Use:
%\interdisplaylinepenalty=2500
% after loading amsmath to restore such page breaks as IEEEtran.cls normally
% does. amsmath.sty is already installed on most LaTeX systems. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/required/amslatex/math/
\usepackage{amssymb}

% *** SPECIALIZED LIST PACKAGES ***
%
%\usepackage{algorithmic}
% algorithmic.sty was written by Peter Williams and Rogerio Brito.
% This package provides an algorithmic environment fo describing algorithms.
% You can use the algorithmic environment in-text or within a figure
% environment to provide for a floating algorithm. Do NOT use the algorithm
% floating environment provided by algorithm.sty (by the same authors) or
% algorithm2e.sty (by Christophe Fiorio) as IEEE does not use dedicated
% algorithm float types and packages that provide these will not provide
% correct IEEE style captions. The latest version and documentation of
% algorithmic.sty can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithms/
% There is also a support site at:
% http://algorithms.berlios.de/index.html
% Also of interest may be the (relatively newer and more customizable)
% algorithmicx.sty package by Szasz Janos:
% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithmicx/

% *** ALIGNMENT PACKAGES ***
%
% \usepackage{array}
% Frank Mittelbach's and David Carlisle's array.sty patches and improves
% the standard LaTeX2e array and tabular environments to provide better
% appearance and additional user controls. As the default LaTeX2e 
% generation code is lacking to the point of almost being broken with
% respect to the quality of the end results, all users are strongly
% advised to use an enhanced (at the very least that provided by array.sty)
% set of table tools. array.sty is already installed on most systems. The
% latest version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/required/tools/

%\usepackage{mdwmath}
%\usepackage{mdwtab}
% Also highly recommended is Mark Wooding's extremely powerful MDW tools,
% especially mdwmath.sty and mdwtab.sty which are used to format equations
% and tables, respectively. The MDWtools set is already installed on most
% LaTeX systems. The lastest version and documentation is available at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/mdwtools/

% IEEEtran contains the IEEEeqnarray family of commands that can be used to
% generate multiline equations as well as matrices, tables, etc., of high
% quality.

%\usepackage{eqparbox}
% Also of notable interest is Scott Pakin's eqparbox package for creating
% (automatically sized) equal width boxes - aka "natural width parboxes".
% Available at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/eqparbox/

% *** SUBFIGURE PACKAGES ***
% \usepackage[tight,footnotesize]{subfigure}
% subfigure.sty was written by Steven Douglas Cochran. This package makes it
% easy to put subfigures in your figures. e.g., "Figure 1a and 1b". For IEEE
% work, it is a good idea to load it with the tight package option to reduce
% the amount of white space around the subfigures. subfigure.sty is already
% installed on most LaTeX systems. The latest version and documentation can
% be obtained at:
% http://www.ctan.org/tex-archive/obsolete/macros/latex/contrib/subfigure/
% subfigure.sty has been superceeded by subfig.sty.
 \ifCLASSOPTIONcompsoc
  \usepackage[tight,normalsize,sf,SF]{subfigure}
\else
  \usepackage[tight,footnotesize]{subfigure}
\fi

%\usepackage[caption=false]{caption}
%\usepackage[font=footnotesize]{subfig}
% subfig.sty, also written by Steven Douglas Cochran, is the modern
% replacement for subfigure.sty. However, subfig.sty requires and
% automatically loads Axel Sommerfeldt's caption.sty which will override
% IEEEtran.cls handling of captions and this will result in nonIEEE style
% figure/table captions. To prevent this problem, be sure and preload
% caption.sty with its "caption=false" package option. This is will preserve
% IEEEtran.cls handing of captions. Version 1.3 (2005/06/28) and later 
% (recommended due to many improvements over 1.2) of subfig.sty supports
% the caption=false option directly:
%\usepackage[caption=false,font=footnotesize]{subfig}
%
% The latest version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/subfig/
% The latest version and documentation of caption.sty can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/caption/

% *** FLOAT PACKAGES ***
%
%\usepackage{fixltx2e}
% fixltx2e, the successor to the earlier fix2col.sty, was written by
% Frank Mittelbach and David Carlisle. This package corrects a few problems
% in the LaTeX2e kernel, the most notable of which is that in current
% LaTeX2e releases, the ordering of single and double column floats is not
% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
% single column figure to be placed prior to an earlier double column
% figure. The latest version and documentation can be found at:
% http://www.ctan.org/tex-archive/macros/latex/base/

\usepackage{stfloats}
% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
% the ability to do double column floats at the bottom of the page as well
% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
% LaTeX2e). It also provides a command:
%\fnbelowfloat
% to enable the placement of footnotes below bottom floats (the standard
% LaTeX2e kernel puts them above bottom floats). This is an invasive package
% which rewrites many portions of the LaTeX2e float routines. It may not work
% with other packages that modify the LaTeX2e float routines. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/sttools/
% Documentation is contained in the stfloats.sty comments as well as in the
% presfull.pdf file. Do not use the stfloats baselinefloat ability as IEEE
% does not allow \baselineskip to stretch. Authors submitting work to the
% IEEE should note that IEEE rarely uses double column equations and
% that authors should try to avoid such use. Do not be tempted to use the
% cuted.sty or midfloat.sty packages (also by Sigitas Tolusis) as IEEE does
% not format its papers in such ways.

%\ifCLASSOPTIONcaptionsoff
%  \usepackage[nomarkers]{endfloat}
% \let\MYoriglatexcaption\caption
% \renewcommand{\caption}[2][\relax]{\MYoriglatexcaption[#2]{#2}}
%\fi
% endfloat.sty was written by James Darrell McCauley and Jeff Goldberg.
% This package may be useful when used in conjunction with IEEEtran.cls'
% captionsoff option. Some IEEE journals/societies require that submissions
% have lists of figures/tables at the end of the paper and that
% figures/tables without any captions are placed on a page by themselves at
% the end of the document. If needed, the draftcls IEEEtran class option or
% \CLASSINPUTbaselinestretch interface can be used to increase the line
% spacing as well. Be sure and use the nomarkers option of endfloat to
% prevent endfloat from "marking" where the figures would have been placed
% in the text. The two hack lines of code above are a slight modification of
% that suggested by in the endfloat docs (section 8.3.1) to ensure that
% the full captions always appear in the list of figures/tables - even if
% the user used the short optional argument of \caption[]{}.
% IEEE papers do not typically make use of \caption[]'s optional argument,
% so this should not be an issue. A similar trick can be used to disable
% captions of packages such as subfig.sty that lack options to turn off
% the subcaptions:
% For subfig.sty:
% \let\MYorigsubfloat\subfloat
% \renewcommand{\subfloat}[2][\relax]{\MYorigsubfloat[]{#2}}
% For subfigure.sty:
% \let\MYorigsubfigure\subfigure
% \renewcommand{\subfigure}[2][\relax]{\MYorigsubfigure[]{#2}}
% However, the above trick will not work if both optional arguments of
% the \subfloat/subfig command are used. Furthermore, there needs to be a
% description of each subfigure *somewhere* and endfloat does not add
% subfigure captions to its list of figures. Thus, the best approach is to
% avoid the use of subfigure captions (many IEEE journals avoid them anyway)
% and instead reference/explain all the subfigures within the main caption.
% The latest version of endfloat.sty and its documentation can obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/endfloat/
%
% The IEEEtran \ifCLASSOPTIONcaptionsoff conditional can also be used
% later in the document, say, to conditionally put the References on a 
% page by themselves.

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
%\usepackage{url}
% url.sty was written by Donald Arseneau. It provides better support for
% handling and breaking URLs. url.sty is already installed on most LaTeX
% systems. The latest version can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/misc/
% Read the url.sty source comments for usage information. Basically,
% \url{my_url_here}.

% *** Do not adjust lengths that control margins, column widths, etc. ***
% *** Do not use packages that alter fonts (such as pslatex).         ***
% There should be no need to do such things with IEEEtran.cls V1.6 and later.
% (Unless specifically asked to do so by the journal or conference you plan
% to submit to, of course. )

% correct bad hyphenation here
% \hyphenation{op-tical net-works semi-conduc-tor}
% Algorithms
\usepackage{float}
\floatstyle{ruled}
\newfloat{algorithm}{htbp}{loa}%[chapter]
\floatname{algorithm}{Algorithm}

@@

\section{State and Parameter Estimation}
The state-space representation allows the use of the expectation maximization (EM) algorithm \cite{Dempster1977,Shumway2000} to infer both the kernel and the field from the electrophysiological data. The EM algorithm is a two-part iterative algorithm that yields the maximum likelihood estimate of the connectivity kernel parameters (M-step) and the posterior distribution of the field weights (E-step). The M-step finds an estimate of $\boldsymbol\theta$ that maximizes the lower bound, $\mathcal{Q}\left(\boldsymbol\theta,\boldsymbol\theta'\right)$, of the complete-data log-likelihood,  $p\left(\mathbf{X},\mathbf{Y} \mid \boldsymbol\theta \right)$. The E-step uses the estimate of $\boldsymbol\theta$, $\boldsymbol\theta'$ and the Rauch Tung Striebel (RTS) smoother \cite{Gibsona2005} to estimate the posterior state sequence. In this paper we use a standard EM methodology. Therefore, we only describe aspects of the algorithm that are specific to MRA neural field equations due to the space limitations. The bound on the log-likelihood is a quadratic in the form of
\begin{equation}
 \mathcal Q\left(\boldsymbol \theta,\boldsymbol\theta'\right)=\beta+2T_s\left(\boldsymbol\upsilon_0-\boldsymbol\upsilon_1\right)\boldsymbol\theta-T_s\boldsymbol\theta^\top\boldsymbol\Upsilon\boldsymbol\theta,
\end{equation}
with a maximum at
\begin{align}
 \boldsymbol \theta= \boldsymbol\Upsilon^{-\top}(\boldsymbol\upsilon_0^\top-\boldsymbol\upsilon_1^\top),
\end{align}
where
\begin{align}\label{eq:upsilon0}
 \boldsymbol\upsilon_0&=\sum_{i,j=1}^{n_x}[\boldsymbol\Xi_0]_{i,j}[\Sigma_{w}^{-1}\boldsymbol\Lambda_{x}^{-1}\mathbf U]_{j,i}\\ 
 \boldsymbol\upsilon_1&=\xi\sum_{i,j=1}^{n_x}[\boldsymbol\Xi_1]_{i,j}[\Sigma_{w}^{-1}\boldsymbol\Lambda_{x}^{-1}\mathbf U]_{j,i}\label{eq:upsilon1} \\
 \boldsymbol\Upsilon&=T_s\sum_{i,j=1}^{n_x}[\boldsymbol\Xi_1]_{i,j}[\mathbf{U}^{\mathsf T} \boldsymbol\Lambda_{x}^{-1}\Sigma_{w}^{-1}\boldsymbol\Lambda_{x}^{-1}\mathbf{U}]_{j,i}\label{eq:Upsilon}.
\end{align}
The $\left(i,j\right)^{th}$ element of the block matrix $\mathbf U_{n_x \times n_x}$ is given by
\begin{align}
\left[ \mathbf U\right] _{i,j}&=\int_{\boldsymbol \Omega}\left[\mu(r) \right]_i \left[\int_{\boldsymbol\Omega} \boldsymbol\mu\left(r'\right)\boldsymbol \lambda^\top \left(\mathbf{r-r'}\right) dr'\right]_{j:} dr.
\end{align}
where $[.]_{j:} $ denotes $j^{th}$ row of the matrix. The second terms in Eqs. \ref{eq:upsilon0}, \ref{eq:upsilon1}, and \ref{eq:Upsilon} can be computed as a once-off before the commencement of the EM iterations, which increases the speed of the M-step significantly compared to the method in \cite{Dewar2009}. $\Xi$-variables are calculated using the smoother outputs, i.e., state estimates, $\hat{\mathbf x}_t$ along with their covariance matrix, $\mathbf P_t$, and the cross-covariance matrix $\mathbf M_t$ using \cite{Gibsona2005}
\begin{align}\label{eq:Xivariables}
\boldsymbol\Xi_0&=\sum_{t=0}^{T-1}\left(\mathbf M_{t+1}+\mathbf{\hat x}_t\mathbf{\hat x}_{t+1}^\top\right) \\
 \boldsymbol\Xi_1&=\sum_{t=0}^{T-1}\left(\mathbf P_t+\mathbf{\hat x}_t\mathbf{\hat x}_t^\top\right).
\end{align}
\dean{Number of iterations and stopping criteria?}The stopping criterion is usually associated  with  either the change in the parameter estimates or the log-likelihood variation \cite{McLachlan1997}.

>>>>

@@

\usepackage{cite}

\ifCLASSINFOpdf
   \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../pdf/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
\fi

\usepackage[cmex10]{amsmath}
\interdisplaylinepenalty=2500
\usepackage{amssymb}
\ifCLASSOPTIONcompsoc
  \usepackage[tight,normalsize,sf,SF]{subfigure}
\else
  \usepackage[tight,footnotesize]{subfigure}
\fi

\usepackage{stfloats}
\usepackage{float}
\floatstyle{ruled}
\newfloat{algorithm}{htbp}{loa}%[chapter]
\floatname{algorithm}{Algorithm}

@@

\section{State and Parameter Estimation}

The state-space representation of the MRAIDE allows the use of the well known Expectation Maximization (EM) algorithm \cite{Dempster1977,Shumway2000} to infer both the kernel and the field from electrophysiological data. The EM algorithm, when used to in this context \cite{Dewar,Scerri,Freestone}, yields the maximum likelihood kernel estimate and the posterior distribution of the field over time. We use the standard EM-based approach for linear dynamic systems \cite{Gibsona2005,Roweis97} \mike{Roweis and Gharamani, 1997, A Unifying Review of Linear Guassian Models} and hence herein only describe aspects of the algorithm that are specific to MRA neural field equations. EM essentially iteratively finds increasingly tighter lower bounds on the likelihood of the kernel so that, at convergence, the maximum of the bound corresponds to the maximum of the likelihood function. For a multivariate linear dynamic system, the bound used is a quadratic of the form
\begin{equation}
 \mathcal Q\left(\boldsymbol \theta,\boldsymbol\theta'\right)=\beta+2T_s\left(\boldsymbol\upsilon_0-\boldsymbol\upsilon_1\right)\boldsymbol\theta-T_s\boldsymbol\theta^\top\boldsymbol\Upsilon\boldsymbol\theta,
\end{equation}
which has a maximum at
\begin{align}
 \boldsymbol \theta= \boldsymbol\Upsilon^{-\top}(\boldsymbol\upsilon_0^\top-\boldsymbol\upsilon_1^\top),
\end{align}
where
\begin{align}\label{eq:upsilon0}
 \boldsymbol\upsilon_0&=\sum_{i,j=1}^{n_x}[\boldsymbol\Xi_0]_{i,j}[\Sigma_{w}^{-1}\boldsymbol\Lambda_{x}^{-1}\mathbf U]_{j,i}\\ 
 \boldsymbol\upsilon_1&=\xi\sum_{i,j=1}^{n_x}[\boldsymbol\Xi_1]_{i,j}[\Sigma_{w}^{-1}\boldsymbol\Lambda_{x}^{-1}\mathbf U]_{j,i}\label{eq:upsilon1} \\
 \boldsymbol\Upsilon&=T_s\sum_{i,j=1}^{n_x}[\boldsymbol\Xi_1]_{i,j}[\mathbf{U}^{\mathsf T} \boldsymbol\Lambda_{x}^{-1}\Sigma_{w}^{-1}\boldsymbol\Lambda_{x}^{-1}\mathbf{U}]_{j,i}\label{eq:Upsilon}.
\end{align}
The $\left(i,j\right)^{th}$ element of the block matrix $\mathbf U_{n_x \times n_x}$ is given by
\begin{align}
\left[ \mathbf U\right] _{i,j}&=\int_{\boldsymbol \Omega}\left[\mu(r) \right]_i \left[\int_{\boldsymbol\Omega} \boldsymbol\mu\left(r'\right)\boldsymbol \lambda^\top \left(\mathbf{r-r'}\right) dr'\right]_{j:} dr.
\end{align}
where $[.]_{j:} $ denotes $j^{th}$ row of the matrix. The second terms in Eqs. \ref{eq:upsilon0}, \ref{eq:upsilon1}, and \ref{eq:Upsilon} can be computed as a once-off before the commencement of the EM iterations, which increases the speed of the M-step significantly compared to the method in \cite{Dewar2009}. $\Xi$-variables are calculated using the smoother outputs, i.e., state estimates, $\hat{\mathbf x}_t$ along with their covariance matrix, $\mathbf P_t$, and the cross-covariance matrix $\mathbf M_t$ using \cite{Gibsona2005}
\begin{align}\label{eq:Xivariables}
\boldsymbol\Xi_0&=\sum_{t=0}^{T-1}\left(\mathbf M_{t+1}+\mathbf{\hat x}_t\mathbf{\hat x}_{t+1}^\top\right) \\
 \boldsymbol\Xi_1&=\sum_{t=0}^{T-1}\left(\mathbf P_t+\mathbf{\hat x}_t\mathbf{\hat x}_t^\top\right).
\end{align}
\dean{Number of iterations and stopping criteria?}The stopping criterion is usually associated  with  either the change in the parameter estimates or the log-likelihood variation \cite{McLachlan1997}.
